I would like to take a moment and extend a special thank you to all of the individuals who have helped me through the journey to complete this thesis. I first would like to thank my adviser, Dr. Christof Teuscher, for all of his guidance throughout my time at Portland State University. Through the ups and downs of this process, Dr. Teuscher always encouraged me to stick with it. I would also like to thank the committee members, Dr. Eric Wan and Dr. Marek Perkowski, for their advice and time on this thesis. I thank to Prof. Mark Faust for the encouragement to opt for the thesis route instead of coursework as well as general advice throughout my career at Portland State. I also extend a thank you to my lab colleague, Peter Banda, that provided me with the help to ramp up on this topic and work with me on publishing my first paper.

I am grateful to my family who encouraged me to complete this program. I appreciate the advice provided throughout this degree by members of my family: Michelle Moles, Steve Moles, Courtney Moles, and Scott Moles. I especially would like to thank my friend, Jacob Couch, for his advice, encouragement, patience, and proofreading. I also would like to acknowledge my friends who have helped me throughout this research and coursework at Portland State. I also appreciate the patience of these outstanding individuals while I completed this thesis while working full time. This includes, and is not limited to Kurt Kolkind, Gabriel Thompson, Daniel Lenski, Brian Jones, Giedre Novikaite, Aaron Anderson, Erik Rhodes, Lindsey Geer, Anne Van Hulle, Eric Krause, and Scott Cline.
